---
title: "R for Humanities: Regular Expressions"
output: 
  html_notebook: 
    number_sections: true
editor_options: 
  markdown: 
    wrap: 72
---

[Main
Menu](https://htmlpreview.github.io/?https://github.com/argentenum/Rworkshop/blob/master/notebooks/RforHumanities.nb.html)

# Regular Expressions

In the previous lesson we looked at one instance of a regular
expressions function. But let us now look at it at greater detail.

## grep

grep gives us the line or vector index where a particular substring
pattern match is found.

```{r}
sonnets <- readLines("https://www.gutenberg.org/files/1041/1041-0.txt")
grep("Shall I compare", sonnets)
```

The result is a vector

```{r}
week
grep("e",week)
```

It gives all the indices where the pattern has found a match.

We can also return the values instead of the indices.

```{r}
grep("e",week, value = TRUE)
```

## regexpr

This tells us if a search string is present in a given string and also gives us the starting position of the search string within the given string.

```{r}
regexpr("uat","Equator")
```
Here it tells us that the search string match was found from the 3rd character of the given string, the search string was 4 characters in length. This only finds the first instance of a match.

## gregexpr 

This functions in a way similar to that of regexpr, but it traces all instances of matches between the search string and the given string.

```{r}
gregexpr("iss", "Mississipi")
```
## Replacing substrings

### sub

replace substring - only first instance 

```{r}
sub("iss","ant","Mississipi")
```

### gsub 
replace substrings - all instances 

```{r}
gsub("iss","ant","Mississipi")
```

# Install 'stringr'

For more advanced regular expressions operations we need the 'stringr' package. The 'stringr' package is part of the 'tidyverse' package. One can also choose to install and call the 'stringr' package.

```{r}
install.packages("stringr")
library(stringr)
```

Let us create a string.

```{r}
basicString <- "Drew has 3 watermelons, Alex has 4 hamburgers, Karina has 12 tamales, and Anna has 6 soft pretzels"
```

## pulling out every instance of one person's name from a string

```{r}
basicExtractAll <- str_extract_all(basicString, "Drew")
basicExtractAll
```
The result is a list.


## replacing pattern with a replacement string

```{r}
basicReplaceAll <- str_replace_all(basicString, "Alex", "Shawn")
basicReplaceAll
```

## matching character set

### here matching vowels 

```{r}
charSets01 <- str_extract_all(basicString,"[aeiou]")
charSets01
```

### matching character range - are case sensitive 

```{r}
charSets02 <- str_extract_all(basicString, "[A-Z]")
charSets02
```


### matching number range 

```{r}
charSets03 <- str_extract_all(basicString, "[0-9]")
charSets03
```


### matching mix of charter types

```{r}
charset04 <- str_extract_all(basicString, "[A-Ct-z1-6]")
charset04
```

## Metacharacters 

### matching spaces 

```{r}
str_extract_all(basicString,"\\s")

```

### matching alphanumeric characters

```{r}
str_extract_all(basicString, "\\w")
```

### matching digits digits

```{r}
str_extract_all(basicString, "\\d")
```

### Quantifiers - specifying how many characters are expected 

```{r}
quantExample <- "B.BB.BBB.BBBB" 
str_extract_all(quantExample, "B+")
```
Here the function matches Bs till end of the group of Bs

```{r}
str_extract_all(quantExample, "B{2}")
```
This matches exactly two Bs till the end of the group of Bs 

```{r}
str_extract_all(quantExample, "B{2,4}")
```
This matches any number from 2 to 4

```{r}
str_extract_all(quantExample, "B{2,}")
```
### Picking up words using quantifier

```{r}
str_extract_all(basicString,"[A-Za-z]+")
```

### find numbers 

```{r}
str_extract_all(basicString,"\\d{1,2}")
```

Find the quantity of each food item ie. each number directly followed by name of the food item separated by space 

```{r}
str_extract_all(basicString, "\\d{1,2}\\s\\w+\\s*\\w*")
```

### Capture groups

```{r}
captureGroup1 <- str_match_all(basicString,"[A-Z][a-z]+\\s\\w+\\s\\d{1,2}\\s\\w+\\s*\\w*")
captureGroup1
```

Leaving out 'has' and preparing data in columns. It is basically the same formula as captureGroup1, just brackets around those strings which are collected as list 

```{r}
captureGroup2 <-str_match_all(basicString,"([A-Z][a-z]+)\\s\\w+\\s(\\d{1,2})\\s(\\w+\\s*\\w*)")
captureGroup2
```

### converting to dataframe 

```{r}
picnic <- data.frame(captureGroup2[[1]][,-1])
colnames(picnic) <- c("Name", "Quantity", "Item")
picnic
```


[Stringr cheatsheet1](https://github.com/rstudio/cheatsheets/blob/main/strings.pdf)

[Stringr cheatsheet2](https://regenerativetoday.com/a-complete-beginners-guide-to-regular-expressions-in-r/)

### Extracting URLs

```{r}
urls <- "Here are urls ONE www.jnu.ac.in, TWO https://arjunghosh.in, THREE http://dharti.wordpress.com"
str_extract_all(urls, "(https?)?(://)?(www\\.)?\\w+\\.[a-zA_Z0-9-.]+")
```

### extract all capitalized words 

```{r}
fewNames <- "My name is Anthony Gonsalves and I am not Amar Akbar Anthony" 
str_extract_all(fewNames, "([A-Z][a-z]+)")[[1]]
```

### two names together 

```{r}
str_extract_all(fewNames,"([A-Z][a-z]+)\\s([A-Z][a-z]+)")
```

### multiple names together

```{r}
str_extract_all(fewNames, "(([A-Z][a-z]+)\\s?){1,}")
```

### strip extra spaces 

```{r}
trimws(str_extract_all(fewNames,"(([A-Z][a-z]+)\\s?){1,}")[[1]])
```


```{r}
king <- "The king of England is Charles Phillip Arthur George, also known as Charles III"
str_extract_all(king, "((([A-Z][a-z]+)\\s*)+){2,}")
```

[Main Menu](https://htmlpreview.github.io/?https://github.com/argentenum/Rworkshop/blob/master/notebooks/RforHumanities.nb.html)
