---
title: "R for Humanities: Map plotting"
output: 
  html_notebook: 
    number_sections: true
editor_options: 
  markdown: 
    wrap: 72
---

[Main Menu](https://htmlpreview.github.io/?https://github.com/argentenum/Rworkshop/blob/master/notebooks/RforHumanities.nb.html)


[Georeferencing](https://map-rfun.library.duke.edu/01_georeference.html)

For plotting points on a map we need to install and run two libraries

```{r}
library(sf)
library(mapview)
```


# reading the data

```{r}
lifeDF <- read.csv("./data/cleanoutput.csv", stringsAsFactors = FALSE)
```

In 'mapview' the coordintes are pushed separately as Longitude and Latitude in decimal format.

# converting data into spatial objects
```{r}
birthLoc <- st_as_sf(lifeDF, coords = c("birthlon","birthlat"),
                     crs=4326)
deathLoc <- st_as_sf(lifeDF, coords = c("deathlon","deathlat"),
                     crs=4326)
```

These are defined as a 'spatial' data type

```{r}
str(birthLoc)
```


# plotting the map

## just one type of location

```{r}
mapview(birthLoc)
```

[View result](https://web.iitd.ac.in/~arjunghosh/files/birthloc.html)

## multiple location types

```{r}
mapview(list(birthLoc,deathLoc),col.regions=list("blue","red"),col=list("blue","red"),label=lifeDF$Name)
```

[View result](https://web.iitd.ac.in/~arjunghosh/files/birdealoc.html)

## Map types
As you can see 'mapview' uses the Leaflet library. [Leaflet](https://leafletjs.com/) is an open-source Javascript library that uses OpenStreetMap as a base layer.

OpenStreetMap provides several views. It gives "CartoDB.Positron" as the default base layer. We can choose to use a different layer.

```{r}
mapview(list(birthLoc,deathLoc),col.regions=list("blue","red"),col=list("blue","red"),label=lifeDF$Name, map.types = "Esri.WorldShadedRelief")
```

[View result](https://web.iitd.ac.in/~arjunghosh/files/birdeadark.html)

## Pop-up labels

In the above maps you get the name of the person as a popup label. That is because we have defined 'label' as a specific vector within the 'lifeDF'.

It is possible to generate an entire table as popup data for each point. This can be done by defining the 'label' as a data.frame

```{r}
mapview(list(birthLoc,deathLoc),col.regions=list("blue","red"),col=list("blue","red"),label=lifeDF, map.types = "Esri.WorldShadedRelief")
```

[View result](https://web.iitd.ac.in/~arjunghosh/files/birdeatable.html)


# Map for analysis

Maps can be used not only as visualization, but also for analysis. Let us demostrate with the following dataset.

```{r}
iitdDF <- read.csv("http://web.iitd.ac.in/~arjunghosh/files/iitdPersons.csv")
head(iitdDF)
```
This is a dataset obtained from wikidata of IIT Delhi alumni and their places of birth.

Lets us create a spatial object and visualize to an interactive map.

```{r}
birthiitd <- st_as_sf(iitdDF, coords = c("Long","Lat"), crs=4326)
mapview(birthiitd, zcol = "yob", label = birthiitd$personLabel, legend = TRUE)
```

Here we are using the 'mapview' function to plot values according to the the Year of Birth (yob).
[View Result](https://web.iitd.ac.in/~arjunghosh/files/iitdmap.html)

As we can see the in the resulting map the points are coloured according to a gradient - old to recent according to the Year of Birth. Using this we can try to ascertain if there has been a change in the places of origin of successful IITD alumni over the years.

# More mapping on R
Though 'mapview' is an easy to use geospatial visualizer, much more is possible using R. Beyond 'mapview', mapping on R can also be undertaken using 'ggplot' and 'ggmaps'. Take a look at this tutorial which advises us on how to create a [Chloropleth map](https://r-graph-gallery.com/183-choropleth-map-with-leaflet.html) or one where we can analyse data available for specific spatial shapes or polygons. This allows us to analyse distribution across geographical categories like districts, states or countries.

[Main Menu](https://htmlpreview.github.io/?https://github.com/argentenum/Rworkshop/blob/master/notebooks/RforHumanities.nb.html)