---
title: "R for Humanities: Lesson 3"
output: 
  html_notebook: 
    number_sections: true
---

[Main Menu](https://htmlpreview.github.io/?https://github.com/argentenum/Rworkshop/blob/master/notebooks/RforHumanities.nb.html)

# Lists

## creating list

```{r}
march <- c(1,2,3,4,5,6,7,8)
week <- c("mon","tue","wed","thu","fri","sat","sun")
marchList <- list(march, week)
marchList
```

lists can have text and string vectors
lists can have vectors of unequal lengths

```{r}
str(marchList)
```


## finding length of list gives the number of columns

```{r}
length(marchList)
```


## use unlist function to merge columns into single vector

```{r}
unlist(marchList)
```


## index a vector

```{r}
week[3]
```

```{r}
week[2:4]
```
the result is also a vector

## index a list

```{r}
marchList[1]
```
the result is a list.

```{r}
str(marchList[1])
```


```{r}
length(marchList)
```

```{r}
length(marchList[1])
```

```{r}
length(unlist(marchList))
```

# Dataframes

## creating dataframe

```{r}
marchFrame <- data.frame(march,week)
```

vectors have to be of equal length to create a data frame

```{r}
marchFrame <- data.frame(march[1:7],week)
```

```{r}
str(marchFrame)
```
The Vector variable/ name becomes the column header.

## indexing dataframe: by column number

```{r}
marchFrame[2]
```

## indexing dataframe: by header name

```{r}
marchFrame$week
```
When indexed by column number the result is a data frame. Basically, it is a subset of the existing data frame.
```{r}
str(marchFrame[2])
```
When indexed by column header the result is a Vector.

```{r}
str(marchFrame$week)
```

## Length of the dataframe gives the number of columns

```{r}
length(marchFrame)
```
```{r}
length(marchFrame[2])
```

```{r}
length(marchFrame$week)
```

## indexing a particular value in a data frame

index by row number and column number

```{r}
marchFrame[4,2]
```

## getting column labels

```{r}
names(marchFrame)
```

the result is a vector

```{r}
str(names(marchFrame))
```
## renaming column labels

```{r}
names(marchFrame) <- c("Order","Days of the Week")
marchFrame
```

## getting and naming column labels

```{r}
rownames(marchFrame)
```
By default they are just numbers as strings

```{r}
rownames(marchFrame) <- c("som","man","bud","gur",
                          "suk","sha","rav")
marchFrame
```

# Count the number of rows and columns in a dataframe

```{r}
nrow(marchFrame)
```
```{r}
ncol(marchFrame)
```


# Exercise
Create a dataframe with the first column giving 
the names of every month in a year and the second 
column giving the corresponding number of days 
in each month.


[Main Menu](https://htmlpreview.github.io/?https://github.com/argentenum/Rworkshop/blob/master/notebooks/RforHumanities.nb.html)