---
title: "R for Humanities: Lesson 5"
output: 
  html_notebook: 
    number_sections: true
---

[View this page in a browser](https://htmlpreview.github.io/?https://github.com/argentenum/Rworkshop/blob/master/notebooks/RforHumanities05.nb.html)

[Go to Lesson 4](https://htmlpreview.github.io/?https://raw.githubusercontent.com/argentenum/Rworkshop/master/notebooks/RforHumanities04.nb.html)

# Logical operations

## ifelse

```{r}
r <- 3
if (r==4) { 
  s <- "yes"
} else {
    s <- "no"
    }
print(s)
```

# Loops

## while loop

Allows the loop to run while a condition is met.
When the condition is not satisfied, the loop breaks.

```{r}
i <- 1
while(i <= 10) i <- i+4
#while loop 2
while(TRUE) {
  i <- i + 4
  if (i > 10) break
}
```


# for loop

```{r}
disco <- c(4,3,2,1)
for (n in disco)
{
  print(n)
}
```


```{r}
for (n in 1:100)
{
  print(n)
}

```

## stepped for loop by sequence

```{r}
for (n in seq(5,100, by = 5))
{
  print(n)
}
```


## Running a loop on all values of a Vector

```{r}
#by length
age <- NULL
for (n in 1:length(disco)) {
  print(disco[n])
  age[n] <- disco[n]*2
}

```


# Importing data: csv

While importing a data from a file, the file may be referenced through a filepath ot a local file, or through an url to a file on the internet

```{r}
BioDF <- read.csv("https://web.iitd.ac.in/~arjunghosh/RworkshopFiles/cleanoutput.csv")
```

The file is read as a dataframe.

```{r}
str(BioDF)
```


# Getting a glimpse of a large dataframe

```{r}
head(BioDF)
```

# Exercise

From BioDF print names of all personalities who were born in an even number year
add a column giving age at death
scan and exercise


# Importing a text file

Let us import text file containing the first paragraph of [Pride and Prejudice]

```{r}
pnpWords <- scan("https://web.iitd.ac.in/~arjunghosh/RworkshopFiles/pnp.txt"," ")
```

This is a Vector

```{r}
str(pnpWords)
```

## Counting the length of each word in the vector

```{r}
pnpWords[1]
```

```{r}
nchar(pnpWords[1])
```

```{r}
nchar(pnpWords)
```

This too is a Vector

## Using a loop to compute the wordlengths to a table

The table to contain two columns: word and character length of each word.
The vector "pnpwords" gives the words.
The character length of each word to be collected in a Vector ("wordLen")

```{r}
wordLen <- NULL #defile the result vector as NULL before initiating the loop
for (n in 1:length(pnpWords)) {
  wordLen[n] <- nchar(pnpWords[n])
}
pnpDF <- data.frame(pnpWords, wordLen, stringsAsFactors = FALSE)
```

```{r}
head(pnpDF)
```

# stringsAsFactors

Did you notice that while preparing "pnpDF" we defined "strinsAsFactors = FALSE"?
What purpose does this serve.

Let us examine closely.

```{r}
str(pnpDF)
```
Here the Vector pnpwords is a clean string vector

Now, if we defined "stringsAsFactors = TRUE" ...

```{r}
pnpDF2 <- data.frame(pnpWords, wordLen, stringsAsFactors = TRUE)
str(pnpDF2)
```

Here you see pnpDF$pnpWords is defined as factors and not characters.
What does this mean? And why can that be a problem?

Let us try to replace a value in pnpDF2$pnpWords

```{r}
pnpDF2$pnpWords[10]
```

```{r}
pnpDF2$pnpWords[10] <- "woman"
```
Thus factors are a distinct class of vectors, different from characters or numerics. If defined as factors they may not be able to function as characters or numerics respectively.

# additional commands

## generating a Vector of unique values in another vector

```{r}
unique(pnpWords)
```

## convert a string to numeric

```{r}
as.numeric("12345")
```


## convert a numeric value to a string

```{r}
as.character(12345)
```

## convert a string value to a date value

```{r}
as.Date("08-03-2022", "%d-%m-%y%y")
```
```{r}
str(as.Date("08-03-2022", "%d-%m-%y%y"))
```

## Converting a numeric value to date

Need to specify the date of origin for date calculations. If not specified the default value is 1 Jan 1970.

```{r}
as.Date(18329, origin = "1970-01-01")
```


## check presence of a substring in a string

```{r}
"truth" %in% pnpWords
```


## remove duplicate data rows from a dataframe

This process is called "deduplication".

At first we identify the duplicate rows and then get rid of them.

```{r}
duplicated(pnpDF)
```

```{r}
which(duplicated(pnpDF))
```

Now to get rid of duplicate rows we subset the inverse of the duplicated indices.

```{r}
pnpDF[!duplicated(pnpDF),]
```


# Frequency counts

## create a table with frequency counts of values in a vector

```{r}
table(pnpWords)
```

The result is of a class "table"

```{r}
str(table(pnpWords))
```

## sorting a table according to the count

```{r}
sort(table(pnpWords))
```

```{r}
sort(table(pnpWords), decreasing = TRUE)
```


[Go to Lesson 6](https://htmlpreview.github.io/?https://raw.githubusercontent.com/argentenum/Rworkshop/master/notebooks/RforHumanities06.nb.html)