---
title: "R for Humanities: Lesson 7"
output: 
  html_notebook: 
    number_sections: true
editor_options: 
  markdown: 
    wrap: 72
---

[Main
Menu](https://htmlpreview.github.io/?https://github.com/argentenum/Rworkshop/blob/master/notebooks/RforHumanities.nb.html)

# Installing packages

Now, let us go beyond R's native functions. As mentioned earlier in R has a wide array of specialized packages to perform specific tasks. Tasks big and small. This helps customize R to a specific task at hand.

These packages are usually written by volunteer coders. R is an Open Source tool. Which means it's source code (the machine language) is available for everyone to modify and improve. This enables community monitoring and community development.

To use a particular package we first need to install the package in our R Studio installation.

This is a one time task for every installation of RStudio. There are two ways in which a package can be installed in R.

To begin with we shall install the package 'dplyr'. 'dplyr' is a package that is equipped to perform sophisticated data transformation, filtering, summarizing.

## GUI installation

+ In the Viewing box at the Bottom Righthand corner of RStudio clicked the "Packages" tab
+ Go to "Packages>Install"
+ In the 'Packages'column search for the package that you want to install. In this case it is "dplyr".
+ Keep all default options including "Install dependencies"
+ Click install.

R will install the package.

## Console installation
Look closely at your Console. The GUI action has resulted in a command.

```{r}
install.packages("dplyr")
```

We can also use this command to install packages.

Usually, since, installation of a package is expected to be a one time process for each RStudio installation. But on R it is not enough to install a package. In each session of R we need to call the package to be able to use its listed functions. This can be done with the 'library' function.


```{r}
library(dplyr)
```

Given the need to call a package before using any of its functions. This can be done by including the 'library' function for each of the packages called as part of a script.

We are now going to turn our attention to the Pipes operation in R.
Pipes can be typically supported through two packages 'dplyr' and 'magrittr'. 
'dplyr' is tailored to work better with dataframes or tidy datasets (neat rectangular datasets).
'magrittr' works better with arithmetic operations. In this lesson I would like to introduce the working of Pipes using simple arithmetic operations before we move to dataframes.

So let us install and call 'magrittr'

```{r}
install.packages("magrittr")
library(magrittr)
```


# Pipes

The mechanism of programming that we have learnt thus far is usually works in a non-linear fashion. Let me explain. 

## Multiline approach
Study the following code where we carry out some basic arithmetic operations:

```{r}

x <- 10 # Start with a number
x <- x * 2 # Multiply by 2
x <- x - 5 # Subtract 5
x <- x^2 # Raise the result to the power of 2
x <- x / 3 # Divide by 3
print(x) # Print the result

```
## Nested approach

```{r}
x <- 10 #assign a value
x <- (((x*2)-5)^2)/3 #nested operation
print(x) #Print the result
```
Both produce the result. But the first approach repeatedly rewrites the value of x in intermediate lines. In very complex operations this is definitely non-optimal. The Nested approach is probably more difficult for the human eye to follow.

## Linear operations using Pipes

With Pipes the operations take a linear shape with the output of one operation directly pushed to the next operation without any intermediate variable being created.

```{r}
x <- 10 %>%
  multiply_by(2) %>%
  subtract(5) %>%
  raise_to_power(2) %>%
  divide_by(3)

print(x) # Print the result
```

In the above code the value of 'x' is pushed through (pipes!) a number of arithmetic operations and ultimately the resulting value is assigned to the original variable - 'x'.

These set of operations have a linear structure. Though - in the above example - each operation is written in separate lines visually, operationally they make a single line of code and the entire chain of operations run as part of a single command.


## Transforming dataframes

Pipes - using the 'dplyr' package are very effective in subsetting and manipulating data in dataframes

Let us work with a built-in R dataset 'iris'

```{r}
head(mtcars)
```


### Filtering Rows with `filter()`

This operation filters rows based on conditions.

```{r}
# Filter cars with 6 cylinders and mpg greater than 20
filtered_data <- mtcars %>%
  filter(cyl == 6, mpg > 20)

print(filtered_data)
```


### Selecting Columns with `select()`

This allows you to select specific columns from a data frame.

```{r}
# Select only 'mpg', 'cyl', and 'hp' columns
selected_columns <- mtcars %>%
  select(mpg, cyl, hp)

print(selected_columns)
```



### Arranging Rows with `arrange()`

Sort rows by specific column values.

```{r}
# Arrange data by mpg in descending order
arranged_data <- mtcars %>%
  arrange(desc(mpg))

print(arranged_data)
```



### Adding New Columns with `mutate()`

Create new columns based on existing ones.

```{r}
# Create a new column 'weight_per_hp' which is weight (wt) divided by horsepower (hp)
mutated_data <- mtcars %>%
  mutate(weight_per_hp = wt / hp)

print(mutated_data)
```



### Summarizing Data with `summarize()`

Summarize data with aggregation functions such as `mean()`, `sum()`, etc.

```{r}
# Calculate the average mpg for the dataset
summary_data <- mtcars %>%
  summarize(mean_mpg = mean(mpg))

print(summary_data)
```



### Grouping Data with `group_by()` and `summarize()`

Group data by a categorical variable and calculate summary statistics for each group.

```{r}
# Group by 'cyl' and calculate the average mpg for each group
grouped_summary <- mtcars %>%
  group_by(cyl) %>%
  summarize(mean_mpg = mean(mpg))

print(grouped_summary)
```



### Chaining Multiple Operations

You can chain multiple `dplyr` operations together to perform more complex transformations.

```{r}
# Filter for cars with mpg greater than 20, select certain columns, and arrange by horsepower
chained_operations <- mtcars %>%
  filter(mpg > 20) %>%
  select(mpg, cyl, hp) %>%
  arrange(desc(hp))

print(chained_operations)
```



### Using `mutate()` with `ifelse()` for Conditional Calculations

You can create conditional columns with `ifelse()` inside `mutate()`.

```{r}
# Add a new column 'efficiency' based on mpg values (high if mpg > 25, otherwise low)
conditional_column <- mtcars %>%
  mutate(efficiency = ifelse(mpg > 25, "high", "low"))

print(conditional_column)
```



### Summarizing Multiple Metrics

You can summarize multiple metrics at the same time.

```{r}
# Calculate both average mpg and horsepower
multi_summary <- mtcars %>%
  summarize(mean_mpg = mean(mpg), mean_hp = mean(hp))

print(multi_summary)
```



### Sampling Rows with `sample_n()` or `sample_frac()`

You can randomly sample a fixed number of rows or a fraction of the data.

```{r}
# Sample 5 random rows from the dataset
sampled_data <- mtcars %>%
  sample_n(5)

print(sampled_data)
```



### Counting Rows with `count()`

Easily count the number of occurrences of a certain group.

```{r}
# Count the number of cars for each value of 'cyl'
cylinder_counts <- mtcars %>%
  count(cyl)

print(cylinder_counts)
```



### Removing Missing Values with `drop_na()` (from `tidyr`)

This operation drops rows with missing values in a specified column.

```{r}
library(tidyr)

# Assuming there are NA values in the dataset (for demonstration purposes)
# Drop rows with missing 'mpg' values
cleaned_data <- mtcars %>%
  drop_na(mpg)

print(cleaned_data)
```



### Pivoting Data with `pivot_longer()` and `pivot_wider()`

You can reshape data from wide to long format or vice versa. For example:

```{r}
# Pivot data to longer format
longer_data <- mtcars %>%
  pivot_longer(cols = c(mpg, hp, wt), names_to = "variable", values_to = "value")

print(longer_data)
```



### Joining Data Frames with `left_join()`

You can join two datasets together based on a common column.

```{r}
# Create two sample data frames
df1 <- data.frame(ID = 1:5, Name = c("Alice", "Bob", "Charlie", "David", "Eva"))
df2 <- data.frame(ID = 3:7, Score = c(89, 95, 76, 88, 92))

# Perform a left join on the 'ID' column
joined_data <- df1 %>%
  left_join(df2, by = "ID")

print(joined_data)
```



These are just some of the many powerful data manipulation functions that `dplyr` provides. By using pipes (`%>%`), you can chain these operations together to build complex transformations in a very readable and efficient way.





[Main
Menu](https://htmlpreview.github.io/?https://github.com/argentenum/Rworkshop/blob/master/notebooks/RforHumanities.nb.html)
