---
title: "R for Humanities: Lesson 4"
output: 
  html_notebook: 
    number_sections: true
---

[Main Menu](https://htmlpreview.github.io/?https://github.com/argentenum/Rworkshop/blob/master/notebooks/RforHumanities.nb.html)


# Appending data to an existing Data Frame

## Add rows

```{r}
marchFrame
```


```{r}
marchFrame <- rbind(marchFrame,c(8,"holiday"))
marchFrame
```

## Adding Columns

```{r}
marchFrame <- cbind(marchFrame,c(11,12,13,14,15,16,
                                 17,18))
names(marchFrame)[3] <- "attendance"
marchFrame
```

# Merging two data frames

Data in two data frames can be merged by a common column

First let us create two data frames with a common data column.
Though the data may have overlap their order may be different.
There may be some mismatch between values of the matching columns. The values may also be present in different order.

```{r}
flowerPrice <- data.frame(c("marigold","petunia","jasmine","rose"),c(20,40,35,150))
names(flowerPrice) <- c("flower","price")
flowerPrice
```

```{r}
flowerColour <- data.frame(c("daffodil","rose","marigold"),c("pink","red","yellow"))
names(flowerColour) <- c("flower","colour")
flowerColour
```

Now merging ...

```{r}
jointFlower <- merge(flowerColour,flowerPrice, by = 'flower')
jointFlower
```

# Subsetting data frames

Subsetting on the basis of a criteria

```{r}
bigAttendance <- subset(marchFrame,marchFrame$attendance>15)
bigAttendance
```

# Exporting a data frame to a csv file.

## More about Working Directory

The working directory can be referenced as "./" in the file path. This allows the user to organise various files in a project in sub-directories such as "./data", "./script", "./output" etc.

## write.csv

```{r}
write.csv(jointFlower, file = "./output/jointFlower.csv")
```


# Concatenating strings

```{r}
paste("this","is","a","vector",sep = " word ")
```

# Getting date and time

```{r}
Sys.Date()
```

```{r}
Sys.time()
```

```{r}
Sys.timezone()
```

# Creating dynamic file name

## Concatenating the time and strings

```{r}
paste("./output/jointFlower",Sys.Date(),".csv", sep = "")
```

## saving csv to dynamic filepath

```{r}
write.csv(jointFlower,
          file = paste("./output/jointFlower",Sys.Date(),".csv", sep = ""))
```


## generating a string by calculation of dynamic values

```{r}
i <- 8
sprintf("the square of %d is %d",i,i^2)
```


[Main Menu](https://htmlpreview.github.io/?https://github.com/argentenum/Rworkshop/blob/master/notebooks/RforHumanities.nb.html)