---
title: "R for Humanities: Wikidata Query"
output: 
  html_notebook: 
    number_sections: true
editor_options: 
  markdown: 
    wrap: 72
---

[Main
Menu](https://htmlpreview.github.io/?https://github.com/argentenum/Rworkshop/blob/master/notebooks/RforHumanities.nb.html)

We can use R to also run Sparql queries from Wikidata (or from any sparql endpoint for that matter)

The easier way of doing this was to copy the R code snippet from the Wikidata Query Builder and use it within R. This would have resulted in a tidy output (csv, dataframe). However, the SPARQL library is no longer supported. Hence, we shall have to use the "httr" library. This will give the result in the form of a list.

```{r}
library(httr)

# Define the Wikidata SPARQL endpoint
endpoint <- "https://query.wikidata.org/sparql"

# Define your SPARQL query
sparql_query <- '
SELECT ?item ?itemLabel WHERE {
  ?item wdt:P31 wd:Q146.  # example query to get items that are instances of "cat" (Q146)
  SERVICE wikibase:label { bd:serviceParam wikibase:language "en". }
} LIMIT 10
'

# Send the query to the endpoint
response <- GET(url = endpoint, query = list(query = sparql_query), add_headers(Accept = "application/sparql-results+json"))

# Parse the response as JSON
content <- content(response, as = "parsed", type = "application/json")

# Extract results
results <- content$results$bindings

str(results)
```
To query this list we have to carefully study the structure of the "results" and call the index.

To get the details related to the first cat in the "results"

```{r}
results[[1]]
```
As we can see that this is a List within a list. The top list has a length of 2.

We can call any of these two lists.

```{r}
results[[1]][[1]]
```

or

```{r}
results[[1]][[2]]
```
If we want to get the name of the cat we can call

```{r}
results[[1]][[2]]$value
```

If we want a vector of all all the cat names we can call

```{r}
sapply(results, function(x) x$itemLabel$value)
```


Now let us use this query technique to gather the names and URIs for all rivers of India.

```{r}
# Define the Wikidata SPARQL endpoint
endpoint <- "https://query.wikidata.org/sparql"

# Define your SPARQL query
sparql_query <- '
SELECT ?river ?riverLabel
WHERE {
  ?river wdt:P31 wd:Q4022 .
  ?river wdt:P17 wd:Q668 .
    
  SERVICE wikibase:label { bd:serviceParam wikibase:language "en". }
}
'

# Send the query to the endpoint
response <- GET(url = endpoint, query = list(query = sparql_query), add_headers(Accept = "application/sparql-results+json"))

# Parse the response as JSON
content <- content(response, as = "parsed", type = "application/json")

# Extract results
results <- content$results$bindings
```

How many results have we obtained?

```{r}
length(results)
```
Let us examine one such result:
```{r}
results[[100]]
```
Lets us now try to obtain a vector of the URIs of all the rivers.

```{r}
URIset <- sapply(results, function(x) x$river$value)
URIset <- paste("wd:",substring(URIset,regexpr("Q",URIset),nchar(URIset)),sep = "")
URIset[1:10]
```

Now let us attempt to capture a whole lot of data about each river.

```{r}
library(httr)

# Define the Wikidata SPARQL endpoint
endpoint <- "https://query.wikidata.org/sparql"

# Define your SPARQL query
sparql_query <- '
SELECT ?river ?riverLabel ?originLabel ?mouthLabel ?tributaryLabel ?discharge ?watershed ?length
WHERE {
  VALUES ?river {wd:Q949183}.
  OPTIONAL {?river wdt:P885 ?origin  .}
  OPTIONAL {?river wdt:P403 ?mouth .}
  OPTIONAL {?river wdt:P974 ?tributary .}
  OPTIONAL {?river wdt:P2225 ?discharge .}
  OPTIONAL {?river wdt:P2053 ?watershed .}
  OPTIONAL {?river wdt:P2043 ?length .}
    
  SERVICE wikibase:label { bd:serviceParam wikibase:language "en". }
}
'

# Send the query to the endpoint
response <- GET(url = endpoint, query = list(query = sparql_query), add_headers(Accept = "application/sparql-results+json"))

# Parse the response as JSON
content <- content(response, as = "parsed", type = "application/json")

# Extract results
results <- content$results$bindings
```

We can coerce the results to a data.frame

```{r}
riversDF <- data.frame(
  river = sapply(results, function(x) x$river$value),
  riverLabel = sapply(results, function(x) x$riverLabel$value),
  discharge = sapply(results, function(x) x$discharge$value),
  watershed = sapply(results, function(x) x$watershed$value),
  rivlength = sapply(results, function(x) x$length$value),
  originLabel = sapply(results, function(x) x$originLabel$value),
  mouthLabel = sapply(results, function(x) x$mouthLabel$value),
  tributaryLabel = sapply(results, function(x) x$tributaryLabel$value),
  stringsAsFactors = FALSE)

riversDF
```

We can now use the vector "URIset" across a loop to extract the dta about multiple rivers.


[Main
Menu](https://htmlpreview.github.io/?https://github.com/argentenum/Rworkshop/blob/master/notebooks/RforHumanities.nb.html)
