---
title: "R for Humanities: Lesson 1"
output: 
  html_notebook: 
    number_sections: true
---

[View this page in a browser](https://htmlpreview.github.io/?https://raw.githubusercontent.com/argentenum/Rworkshop/master/notebooks/RforHumanities01.nb.html)

R is a a very powerful programming language with a focus on statistical analysis (ie. structured data). It is also able to produce very neat and customisable charts. R is Open Source and this means it is free and 'free'. And that there are many developers who are constantly working voluntarily to improve the tools for R.

A first installation of R comes with a set of standard packages and commands. However, beyond this R has almost 10,000 packages that can be used for various programming task. These packages are mostly hosted on CRAN and can be installed through RStudio. Given these packages R can perform a wide range of tasks that are relevant for humanities scholars (ie. unstructured data).

# Getting started: Installation

For the purpose of this tutorial we will work on RStudio

Depending on the system that you are working on you need to ensure that you have installed R as well as RStudio. If you are working on a Linux system, in all probability, your computer comes equipped with R. But you would still need to install RStudio.

You can find a lot of help for installation of R and RStudio. Here is one https://techvidvan.com/tutorials/install-r/ 

Once installed, launch RStudio and study the interface.

# RStudio Interface

https://hbctraining.github.io/Intro-to-R/img/Rstudio_interface.png

The RStudio interface has four main panels:

**Console**: where you can type commands and see output. The console is all you would see if you ran R in the command line without RStudio.
    
**Script editor**: where you can type out commands and save to file. You can also submit the commands to run in the console.
    
**Environment/History**: environment shows all active objects and history keeps track of all commands run in console
    
**Files/Plots/Packages/Help**


# Working directory

## Viewing your working directory

Before we organize our working directory, let’s check to see where our current working directory is located by typing into the console:

```{r}
getwd()
```

## Running code on R

To run a single line of code you can place the cursor in that line and click Ctrl+Enter
To run a chuck of code, highlight that chunck (set of lines) and click Ctrl+Enter
To run the entire code you can click the "Run" button in the Script editor.


## Setting working directory
Your installation will have a default working directory for R. But you may want to point your R session to a particular working directory or folder so that all scripts and output gets saved at a convenient location. You can in two ways.

You can do it through the GUI - Session > Set working directory > Choose directory

The working directory will get reset each time you start a new session. If you want that a script should run in a particular directory you can write a line of code to point to a particular directory on your computer.

```{r}
setwd("C:/Users/HP/Downloads/Rnotebook")
```

Make sure that you set the url of your choice.

Now we are ready to start our first R operations

# Basic operations

## simple arithmetic operations

```{r}
1+1
```


## commenting

## print function
```{r}
print(1+1)
```


## invoking help

type "?" followed by the function in the Console. The Help topic will be displayed in the bottom right-hand box.

```{r}
?print
```

## printing strings
```{r}
print("workshop")
```


# Variables

## invoking and assigning variable and value

```{r}
a <- 1000
```


##reassigning value to a variable

```{r}
a <- 2000
```


## printing variables

```{r}
print(a)
```


## assigning strings to variables

```{r}
b <- "workshop"
```


## variables can be alphanumeric

```{r}
hijibiji300 <- "workshop"
print(hijibiji300)
```


## variables need to be contiguous

```{r}
an apple <- 23 #produces errors
```

```{r}
getApples <- 23
getApples
```


## use meaningful variables

```{r}
getBananas <- "nuts"

```


## study structure of a value

```{r}
str(getBananas)
```


[Go to Lesson 2](https://htmlpreview.github.io/?https://github.com/argentenum/Rworkshop/blob/master/notebooks/RforHumanities02.html)
