---
title: "R for Humanities: Sentiment Analysis"
output: html_notebook
---

[Main Menu](https://htmlpreview.github.io/?https://github.com/argentenum/Rworkshop/blob/master/notebooks/RforHumanities.nb.html)

In this lesson we shall learn preliminary steps in conducting Sentiment Analysis of texts.

In order to do so we need to learn how to define custom functions in R.

# generating custom functions

In R we can create a macro by defining a series of operations to a function variable.

```{r}
funcofx <- function(x){x+1}
```

Here we are defining a custom function 'funcofx' as any function of a numeric value 'x' that will then peform the arithmetic operation 'x+1'.

Once defined 'funcofx' will perform the artimetic operation over any value of x that is supplied to it.

```{r}
funcofx(10)
```
Such functions can be defined for string operations as well.

```{r}
funcofy <- function(y){
  paste(y, "is the phrase", sep = " ")
}
funcofy("expecto petronum")
```


Now that we have learnt how to define a function let us move on to Sentiment Analysis. We need the package 'sentimentr'

```{r}
library(sentimentr)
```

It gives us a score when we push a string to it.

```{r}
sentiment("The queen is dead. Long live the king")
```

We can see here that the function 'sentiment' recognizes that the string consists of two sentences. It assigns a sentiment score of -0.5 for the first sentence and 0.05 for the second sentence. We can interpret this on a scale of -1 to +1 from negative to positive emotion. Thus the first sentence is assigned a strong negative value on emotions. The second sentence is largely neutral emotionally if only mildly positive.

If we want a deep dive into the words that are producing the sentiment value as assigned by 'sentimentr' we can do so by extracting the sentiment terms.

```{r}
extract_sentiment_terms("The queen is dead. Long live the king")
```

In the console this function produces the following result:

Key: <element_id, sentence_id>
   element_id sentence_id negative positive
        <int>       <int>   <list>   <list>
1:          1           1     dead         
2:          1           2              king

Beyond the assignment of sentiment value within the positive and negative range, we can also extract deeper level of emotions from the string.

```{r}
emotion("The queen is dead. Long live the king")
```

Now let us work with a larger [dataset](https://github.com/argentenum/Rworkshop/blob/master/data/corpus/1836MrPickwickChristmas.txt)

```{r}
sampleText <- readLines("./data/corpus/1836MrPickwickChristmas.txt")
```


# plotting emotions

With series of statements we can plot the emotions.

```{r}
senti <- sentiment(get_sentences(sampleText))
senti <- subset(senti,senti$sentiment!=0) #get rid of emotions non-existent in the text
densitySentiments <- density(senti$sentiment) #get statistics on the sentiments
plot(densitySentiments,main='Density of sentiments')
polygon(densitySentiments,col='dark green')
```


# making a function of the above steps

```{r}
makeSentiPlot <- function(txt){
  senti <- sentiment(get_sentences(txt))
  senti <- subset(senti,senti$sentiment!=0)
  densitySentiments <- density(senti$sentiment) 
  plot(densitySentiments,main='Density of sentiments')
  polygon(densitySentiments,col='dark green')
}
makeSentiPlot(sampleText)
```


# plotting emotions

```{r}
emoti <- emotion_by(get_sentences(sampleText), drop.unused.emotions=TRUE)
plot(emoti) #gives variations of individual emotions in a given piece of text
```


# make a barplot of different emotions 

```{r}
library(dplyr)
```


```{r}

sampleText %>%
  get_sentences() %>%
  emotion_by(drop.unused.emotions=TRUE) %>%
  group_by(emotion_type)  %>%
  summarise(ave_emotion=mean(ave_emotion)) -> txtSummary
barplot(txtSummary$ave_emotion,
        names.arg = txtSummary$emotion_type, las=2, col='red')
```


[Main Menu](https://htmlpreview.github.io/?https://github.com/argentenum/Rworkshop/blob/master/notebooks/RforHumanities.nb.html)