---
title: "R for Humanities: Lesson 2"
output: 
  html_notebook: 
    number_sections: true

---

[View this page in a browser](https://htmlpreview.github.io/?https://github.com/argentenum/Rworkshop/blob/master/notebooks/RforHumanities02.html)

[Go to Lesson 1](https://htmlpreview.github.io/?https://raw.githubusercontent.com/argentenum/Rworkshop/master/notebooks/RforHumanities01.nb.html)

# Vectors

## Numerical vectors
```{r}
march <- c(1,2,3,4,5,6,7,8)
str(march)
print(march)
```


## String vectors

```{r}
week <- c("mon","tue","wed","thu","fri","sat","sun")
str(week)
week
```


## Vector values can either numeric or string

```{r}
mismatch <- c(1,"one")
print(mismatch)
```


## sort function

```{r}
sort(week)
```

## sort in descending order

```{r}
sort(week, decreasing = TRUE)
```

# Various methods of creating vectors

## Specifying the vector elements
```{r}
x <- c(1,3,5)
x
```
## Specifying a range
```{r}
y<- c(1:5)
y
```

## Adding values to a vector

```{r}
x <- c(x,6,7)
x
```

## Append value in between vector values

```{r}
x <- append(x,8,2)
x
z <- append(x,c(100,200),5)
z
```

## Appending character value to numerical vector

```{r}
x <- append(x, "bad")
x
```

## Creating vector through sequence

```{r}
y <- seq(1,100, by = 5)
y
```
```{r}
y <- seq(1,100, by = (3*3))
y
```


## Creating vector by division

```{r}
f <- seq(1,100, length= 25)
f
```
```{r}
f <- seq(1,50, length=6)
f
```

# Vector indices

## Fetching values by index

```{r}
y
z <- c(1,3,5)
y[z]
```
## Fetching inverse indices

```{r}
y[-1]
y[-z]
```

# Vector arithmetic

## Vector arithmetic with vectors of equal lengths

```{r}
u <- c(2,4,6)
u
z
```
```{r}
z+u
```
```{r}
u-z
```
```{r}
u*z
```
```{r}
u/z
```
## Vector arithmetic with vectors of unequal lengths

```{r}
t <- c(1,2,3,4)
t
```
```{r}
t+u
```

```{r}
u-t
```

```{r}
u*t
```

```{r}
t*u
```

```{r}
u/t
```

# The Selection Function - "which"

## Logical symbols 
==  equal to
>   greater than
<   less than
>=  greater than or equal to
<=  less than of equal to
!=  not equal to

```{r}
Symbol <- c("==",">","<",">=","<=","!=")
  Statement <- c("equal to","greater than","less than","greater than equal to","less than equal to","not equal to")
knitr::kable(
  data.frame(Symbol,Statement),
  caption = 'Logical symbols in R'
)
```

## Returning vector indices on condition

```{r}
y>50
```

## 'which' function - logical

```{r}
which(y>50)
```


## returning values that satisfy condition

```{r}
y[which(y>50)]
```
```{r}
week
which(week == "thu")
```


## Test for Divisibility

```{r}
y
y%%2
```


## finding even values

```{r}
which(y%%2==0)
```

```{r}
y[which(y%%2==0)]
```


## any and all

```{r}
any(y>50)
```

```{r}
all(y>50)
```


## if else

```{r}
ifelse(y >50, "yes","no")
```



